---
title: "Bronx Residential Sales August-July"
author: "Blandon Casenave"
date: "November 18, 2015"
output: html_document
---
The Histogram below show us sales volume by tax class in the Bronx, New York. Class 1 = 1-3 Unit Residential Property, Class 2 = Residential Units of 3 or more Units, Class 3 = Utility Company and Class 4 = All Other Properties.

The overwhelming majority of sales volume in the Bronx is generated by residential properties with 1-3 units. 

```{r,echo=FALSE}
TaxClass <- read.csv("https://raw.githubusercontent.com/Misterresearch/CUNY-Projects/1cff0b7218c3138cc5867de2b453191c2b12a605/TaxClass.csv")
hist (TaxClass$TAX.CLASS, breaks = 8, main = "Bronx Sales Building Tax Class", xlab = "Class No.", border = "blue", col = "grey")
```

The following box plot depicts the median values that residential sales prices in the Bronx, NY cluster around.

```{r,echo=FALSE}
SalesPrice <- read.csv("https://raw.githubusercontent.com/Misterresearch/CUNY-Projects/1cff0b7218c3138cc5867de2b453191c2b12a605/SalesPrice.csv")
options(scipen = 10)
boxplot (SalesPrice$SALE.PRICE, main = "Bronx Sales Prices", border = "blue", ylim = c(0,40000000))
```

The following scatter plot depicts the relationship between Sales Prices and Gross Square Footage

```{r,echo=FALSE}
FootageSalesPrice <- read.csv("https://raw.githubusercontent.com/Misterresearch/CUNY-Projects/1cff0b7218c3138cc5867de2b453191c2b12a605/FootageSalesPrice.csv")

options(scipen = 10)
plot (FootageSalesPrice$SQUARE.FEET, FootageSalesPrice$SALE.PRICE, col = "red", xlab = "Gross Sq. Feet", ylab = "Sales Price", ylim = c(0,40000000), xlim = c(0,400000))
```

Source: <a href=http://www1.nyc.gov/site/finance/taxes/property-rolling-sales-data.page?>NYC Finance Rolling Sales Aug.'14 - Jul.'15</a>.  