---
title: "Cross-Platform Video Prod"
author: "Blandon Casenave"
date: "May 19, 2016"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Video Views Section
# The fill argument was required for both of the data sets below, the first time I encountered the error scan message.

AppVideoURL <- "https://raw.githubusercontent.com/Misterresearch/CUNY-Projects/master/ottmobile.csv"

AppVideo <- read.csv(file = AppVideoURL, header = TRUE, sep = ",", strip.white = TRUE, na.strings = "", fill = TRUE)

#remove extraneous observations n: 2371 -> 772.
AppVideo <- AppVideo[AppVideo$FEP.Starts != "0",]

#stringAsFactors set to false doesn't seem to convert vars to characters.
AppVideo$Platform = as.character(AppVideo$Platform)
AppVideo$Network = as.character(AppVideo$Network)
AppVideo$Network<-tolower(AppVideo$Network)
AppVideo$Program = as.character(AppVideo$Program)
#discovered iconv, handy for web based data sources.
AppVideo$Program <- iconv(AppVideo$Program,"WINDOWS-1252","UTF-8")
AppVideo$Program<-tolower(AppVideo$Program)
AppVideo$FEP.Starts = as.numeric(AppVideo$FEP.Starts)
#Recoding work done here, tranformed character values, used levels as reference.
#levels(AppVideo$Network)
AppVideo$Network[AppVideo$Network=="nbcentertainment"] <- "NBC"
AppVideo$Network[AppVideo$Network=="Telemundo Now"] <- "Telemundo"


#App Video Data Frame
AppVideoPlatform<-AppVideo$Platform
AppVideoNetwork <- AppVideo$Network
AppVideoProgram <- AppVideo$Program
AppVideoStarts <- AppVideo$FEP.Starts
AppVideoDF <- data.frame(AppVideoPlatform, AppVideoNetwork, AppVideoProgram, AppVideoStarts, row.names = NULL, check.rows = FALSE, stringsAsFactors = FALSE)


BrowserVideoURL <- "https://raw.githubusercontent.com/Misterresearch/CUNY-Projects/master/browser.csv"
BrowserVideo <- read.csv(file = BrowserVideoURL, header = TRUE, sep = ",", strip.white = TRUE, na.strings = "", fill = TRUE)

#remove extraneous observations, n: 8331 -> 571.
BrowserVideo <- BrowserVideo[BrowserVideo$FEP.Starts >= "2",]

#stringAsFactors set to false doesn't seem to convert vars to characters.
BrowserVideo$Platform = as.character(BrowserVideo$Platform)
BrowserVideo$Network = as.character(BrowserVideo$Network)
BrowserVideo$Network<-tolower(BrowserVideo$Network)
BrowserVideo$Program = as.character(BrowserVideo$Program)
#discovered iconv, handy for web based data sources.
BrowserVideo$Program <- iconv(BrowserVideo$Program,"WINDOWS-1252","UTF-8")
BrowserVideo$Program<-tolower(BrowserVideo$Program)
BrowserVideo$FEP.Starts = as.numeric(BrowserVideo$FEP.Starts)

#Recoding work done here, tranformed character values, used levels as reference.
#levels(BrowserVideo$Network)
BrowserVideo$Network[BrowserVideo$Network=="e"] <- "E!"
BrowserVideo$Network[BrowserVideo$Network=="mun2"] <- "NBC Universo"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - Bravo"] <- "Bravo"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - CNBC"] <- "CNBC"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - E!"] <- "E!"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - Esquire"] <- "Esquire"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - MSNBC"] <- "MSNBC"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - NBC"] <- "NBC"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - NBC News"] <- "NBC News"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - NBC Universo"] <- "NBC Universo"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - Oxygen"] <- "Oxygen"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - Sports Network"] <- "Sports Network"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - Sprout"] <- "Sprout"
BrowserVideo$Network[BrowserVideo$Network=="NBCU TVE - Syfy"] <- "Syfy"

BrowserVideo$Platform[BrowserVideo$Platform=="Mobile"] <- "HTML5"

#Browser Video Data Frame
BrowserVideoPlatform <-BrowserVideo$Platform
BrowserVideoNetwork <- BrowserVideo$Network
BrowserVideoProgram <- BrowserVideo$Program
BrowserVideoStarts <- BrowserVideo$FEP.Starts
BrowserVideoDF <- data.frame(BrowserVideoPlatform,BrowserVideoNetwork, BrowserVideoProgram, BrowserVideoStarts, stringsAsFactors = FALSE)

#Network grouping using base R, dplyring functions not working.
Browsergroup <-(aggregate(BrowserVideoStarts ~ BrowserVideoNetwork, BrowserVideoDF, sum))
Appgroup <-(aggregate(AppVideoStarts ~ AppVideoNetwork, AppVideoDF, sum))
Browsergroup <- Browsergroup[order(-Browsergroup$BrowserVideoStarts),]
Appgroup <-Appgroup[order(-Appgroup$AppVideoStarts),]


Videomerge <- merge(x=Browsergroup, y=Appgroup, by.x = 'BrowserVideoNetwork', by.y = 'AppVideoNetwork')
AllVideoStarts <-(Videomerge$BrowserVideoStarts +Videomerge$AppVideoStarts)
AllVideoNetworks <-Videomerge$BrowserVideoNetwork
AllVideoDF <- data.frame(AllVideoNetworks, AllVideoStarts, stringsAsFactors = FALSE)
AllVideoDF <-(aggregate(AllVideoStarts ~ AllVideoNetworks, AllVideoDF, sum))
AllVideoDF <-AllVideoDF[order(-AllVideoDF$AllVideoStarts),]
#Browsergroup
#Appgroup
#AllVideoDF

#Platform grouping using base R
Browserplat <-(aggregate(BrowserVideoStarts ~ BrowserVideoPlatform, BrowserVideoDF, sum))
Appplat <-(aggregate(AppVideoStarts ~ AppVideoPlatform, AppVideoDF, sum))
Browserplat <- Browserplat[order(-Browserplat$BrowserVideoStarts),]
Appplat <-Appplat[order(-Appplat$AppVideoStarts),]

Videoplat <-merge(x=Appplat, y=Browserplat, by.x='AppVideoPlatform', by.y='BrowserVideoPlatform', all = TRUE)
Videoplat$AppVideoStarts<-ifelse(is.na(Videoplat$AppVideoStarts), 0, Videoplat$AppVideoStarts)
Videoplat$BrowserVideoStarts<-ifelse(is.na(Videoplat$BrowserVideoStarts), 0, Videoplat$BrowserVideoStarts)
AllPlatStarts <-(Videoplat$BrowserVideoStarts + Videoplat$AppVideoStarts)
AllPlat <-(Videoplat$AppVideoPlatform)
AllPlatDF <- data.frame(AllPlat, AllPlatStarts, stringsAsFactors = FALSE)
AllPlatDF <-AllPlatDF[order(-AllPlatDF$AllPlatStarts),]

#Top 25 Program Data Frame
Browserprogram <-(aggregate(BrowserVideoStarts ~ BrowserVideoProgram, BrowserVideoDF, sum))
Appprogram <-(aggregate(AppVideoStarts ~ AppVideoProgram, AppVideoDF, sum))
Browserprogram <- Browserprogram[order(-Browserprogram$BrowserVideoStarts),]
Appprogram <-Appprogram[order(-Appprogram$AppVideoStarts),]

Videoprogram <-merge(x=Appprogram, y=Browserprogram, by.x='AppVideoProgram', by.y='BrowserVideoProgram', all = TRUE)
Videoprogram$AppVideoStarts<-ifelse(is.na(Videoprogram$AppVideoStarts), 0, Videoprogram$AppVideoStarts)
Videoprogram$BrowserVideoStarts<-ifelse(is.na(Videoprogram$BrowserVideoStarts), 0, Videoprogram$BrowserVideoStarts)
AllProgStarts <-(Videoprogram$BrowserVideoStarts + Videoprogram$AppVideoStarts)
AllProg <-(Videoprogram$AppVideoProgram)
AllProgDF <- data.frame(AllProg, AllProgStarts, stringsAsFactors = FALSE)
AllProgDF <-AllProgDF[order(-AllProgDF$AllProgStarts),]

```


With a portfolio of brands, NBCUniversal has a need to understand the performance of each business in relationship to each other.  Within its Entertainment vertical, full-episode video viewing on its ad supported digital platforms is most crucial to the business. 

Below is a bar chart, ranked in descending order of video views across all digital platforms by brand.  Here we can see that NBC Broadcast, as expected has by far the most video views of NBCU's digital entertainment brand. However, among the cable properties Bravo is the strongest network.  We can also see from this chart, the impact of the Hispanic market, which has pushed Telemundo into the third most viewed network within the NBCU portfolio - placing ahead of general market brands such as the USA Network, Syfy & Oxygen.  Although NBCU's news brands CNBC & MSNBC are listed here, they are not expected to perform in this catergory because of most of their viewing is driven by live events and short-form news clips. 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Barplots for each data set
barplot(AllVideoDF$AllVideoStarts, col = "gray", main = "All Views by Network", xlab = "Video Starts", border = "blue", horiz = FALSE, beside = TRUE, names.arg = AllVideoDF$AllVideoNetworks, cex.names = .6, cex.axis = .6, las=2)
```

Given that many of the brands share the same product stack (CMS, Video Player, Mobile Applications), in as much as it is important to understand performance by brand; it's also important to understand performance by platform because that's how investments are made.  It's clear from the bar chart below that we are in midst of another transformative phase in Digital usage, especially for the consumption of video content.  The bar chart below shows, that inspite of the smaller screen people are conusming more full-episodic video on their mobile devices with an app (although still in teh home), than on PC/Laptops. However, we can also see a substantial share of traffic going to connected-devices (AppleTV, Roku, Xbox One, Chromecast, etc.).  The question becomes, what will be the device/platform of the future and where should content providers invest most of their time and resources.  For example, Q12016 was the first time that Apple reported quarterly losses, <a href=https://www.theguardian.com/technology/2016/apr/26/apple-iphone-first-revenue-decline-13-years/>driven by a 16% decrease</a> in the number of iPhones that were sold (which represent two-thirds of its revenue).  Are Apple's lossed the by-product of competition from other mobile phone operating systems, or a shift to other devices? 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
barplot(AllPlatDF$AllPlatStarts, col = "blue", main = "All Views by Platform", xlab = "Video Starts", border = "red", horiz = FALSE, beside = TRUE, names.arg = AllPlatDF$AllPlat, cex.names = .6, cex.axis = .6, las=2)
```

Finally, a look at programming is key determining strategy for any content provider.  After all, it is the programming that provides the incentive for the consumer to visit content providers websites and applications. In this case, consistent with our first chart, nine out of the top ten programs are driven by NBC Broadcast, with number ten belonging to Bravo (the number two network).

In part II of our project, we will exam text found in our Twitter API search for the number program "The Blacklist", and determine if we can glean any insights from it. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
barplot(head(AllProgDF$AllProgStarts, n=10), col = "red", main = "Top 10 Programs", xlab = "Video Starts", border = "blue", horiz = FALSE, beside = TRUE, names.arg = head(AllProgDF$AllProg, n=10), cex.names = .6, cex.axis = .6, las=2)
```

<b>Function Reference Section:</b>

fill argument to help read.table execute<br>
Source: <a href=http://stackoverflow.com/questions/18161009/error-in-reading-in-data-set-in-r>Stack Overflow Posting on Error in Reading in Data Set</a>

iconv function to handle bad characters<br>
Source: <a href=http://stackoverflow.com/questions/13187605/error-in-tolower-invalid-multibyte-string>Stack Overflow Posting on Invalin Multibyte String</a>

full join using base merge<br>
Source: <a href=http://www.r-bloggers.com/joining-data-frames-in-r/>R-Bloggers</a>

addtional cleaning of NA
Source: <a href=http://stackoverflow.com/questions/2613420/handling-missing-incomplete-data-in-r-is-there-function-to-mask-but-not-remove>Stack Overflow</a>