---
title: "DATA606Casenave - Lab5"
author: "Blandon Casenave"
date: "10/18/2016"
output: pdf_document
---
-------PRACTICE-----
```{r, echo=FALSE, fig.align='center', warning=FALSE, eval=TRUE}
library(IS606)
load("more/nc.RData")

summary(nc)
boxplot(nc$weight, nc$habit)
by(nc$weight, nc$habit, mean)

inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")

inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```
-------PRACTICE-----

-------GRADED EXERCISES-------

#1
```{r, echo=FALSE, fig.align='center', warning=FALSE, eval=TRUE}
inference(y = nc$weeks,  est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

#2
```{r, echo=FALSE, fig.align='center', warning=FALSE, eval=TRUE}
inference(y = nc$weeks,  est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conflevel = .90)
```

#3
```{r, echo=FALSE, fig.align='center', warning=FALSE, eval=TRUE}
inference(y = nc$weight,  x=nc$mature, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

#4
Use min/max age for each segment to determine boundaries. Initially my plan was to use assume normally distributed and use 3xSD to determine boundaries...but there was overlap. Also, key word is "non-inference" for this problem. The method of using min/max yields 13-34 for younger moms, and 35-50 for mature moms.
```{r, echo=FALSE, fig.align='center', warning=FALSE, eval=TRUE}
maturem <- subset(nc, mature=="mature mom", select=mage)
youngm <- subset(nc, mature=="younger mom", select=mage)
summary(maturem)
summary(youngm)
```

#5 - Relationship Between Birth Weight & Premie

Based on diffrence between mean birth weights of premature and non-prematurely born babies,their standard error and p-values; we can conclude there is a statisical difference between the average birth weights of the two groups. 

```{r, echo=FALSE, fig.align='center', warning=FALSE, eval=TRUE}
inference(y = nc$weight, x = nc$premie, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")
```