---
title: "DATA606 - Lab2"
author: "Blandon Casenave"
date: "September 9, 2016"
output: pdf_document
---
INCLUDES PRACTICE EXERCISES 
```{r, echo=TRUE, warning=FALSE,fig.align='center'}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
head(kobe)
#Streaks aren't apparent here 
kobe$basket[1:9]
#Kope streak data
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
mean(kobe_streak)
max(kobe_streak)

#simulating independent shooter
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(.45, .55))
hit <- subset(sim_basket, sim_basket == "H")

kobehit <- subset(kobe$basket, kobe$basket == "H")
length(hit)
length(kobehit)
```
INCLUDES PRACTICE EXERCISES

Q1 - Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? 

How long is the player’s longest streak of baskets in 133 shots? sim_basket, with a distribution set to a 45% hit rate, means that most of the shots are actually misses which result in a value of "0". In my first sim run, the typical streak has a mean value of .71 (median = 0), with the longest streak at 5 (see code below). Note: one first simulation yielded results that were almost an exact match with Kobe's stats.
```{r, echo=TRUE, warning=FALSE,fig.align='center'}
sim_streak <- calc_streak(sim_basket)
barplot(table(sim_streak))
mean(sim_streak)
max(sim_streak)
```

Q2- If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

As eluded to above, my results on subsequent runs of sim shots were different. I do not expect the results to be the same, because each simulation is independent of the last one. I do expect the average values to be somewhat close, but I expect the max streaks to vary significantly with each simulation run, as the data has shown.  What I expect to be close are the total counts of hits out of the same number of attempts for each run...because the ratio of Hits to shots still has to be around 45%. 

Q3 - How does Kobe Bryant’s distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe’s shooting patterns? Explain.

The streak lengths are close, most of the time. Kobe's average streak is .76, and it's close to the simutlates shooter's mean which hovers betweem .7-.9 most of the time.  However, I've seen the simulated max range from 4-8, while Kobe's max is four.  This indicates that Kobe's shooting is not a hot-hand, because it's the same or less than that of the simulated shooter whose results are indeed independent of each other. (see first code chunk for Kobe's results) 
