---
title: "DATA606Casenave - Lab4b"
author: "Blandon Casenave"
date: "10/8/2016"
output: pdf_document
---

<br>----------- PRACTICE EXERCISE ONLY-----------<br/>
```{r, echo=TRUE,warning=FALSE, fig.align='center'}
amesURL <- "https://raw.githubusercontent.com/Misterresearch/CUNY-Projects/master/ames2000_NAfix.csv"
ames <- read.csv(amesURL, header = TRUE, sep = ",")
load("more/ames.RData")
#Exercise 1
population <- ames$Gr.Liv.Area
samp <- sample(population, 60)
sample_mean <- mean(samp)
summary(samp)


#R Code for SE, lower/upper CI Limit
se <- sd(samp) / sqrt(60)
lower <- sample_mean - 1.96 * se
upper <- sample_mean + 1.96 * se
c(lower, upper)

#Create empty vector
samp_mean <- rep(NA, 50)
samp_sd <- rep(NA, 50)
n <- 60

#Create loop many sample means and sd
for(i in 1:50){
  samp <- sample(population, n) # obtain a sample of size n = 60 from the population
  samp_mean[i] <- mean(samp)    # save sample mean in ith element of samp_mean
  samp_sd[i] <- sd(samp)        # save sample sd in ith element of samp_sd
}

#CI vector 
#Lastly, we construct the confidence intervals.

lower_vector <- samp_mean - 1.96 * samp_sd / sqrt(n) 
upper_vector <- samp_mean + 1.96 * samp_sd / sqrt(n)

c(lower_vector[1], upper_vector[1])
```


<br>----------------- LAB 4b PROBlEMS BELOW------------------<br/>

Q1 - The plot below shows how instances of the sampled mean are more than two standard deviations (1.96) from the mean, where as the CI, say 95% refers to how many observartions are more than 1.96 deviations from the mean. These numbers will not match because we're not measuring the exact same thing.  So in this case 2 of the 50 sampled means are outside the range or 96%, approximately, but not the exact same as our 95%.  
```{r, echo=TRUE,warning=FALSE, fig.align='center'}
plot_ci(lower_vector, upper_vector, mean(population))
```

Q2 - The Z critical value for a 99% CI is 2.58

Q3 - See the output below, similar to Q1, we see that 1 out of the 50 sampled means values outside of the range or 98% - almost but not the exact same as our 99% CI. 
```{r, echo=TRUE,warning=FALSE, fig.align='center'}
lower_vectorA <- samp_mean - 2.58 * samp_sd / sqrt(n) 
upper_vectorA <- samp_mean + 2.58 * samp_sd / sqrt(n)

c(lower_vectorA[1], upper_vectorA[1])
plot_ci(lower_vectorA, upper_vectorA, mean(population))
```