---
title: "DATA606 - Lab3"
author: "Blandon Casenave"
date: "September 21, 2016"
output: pdf_document
---

---PRACTICE---
```{r, echo=FALSE, fig.align='center'}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")
head(bdims)
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)

#Men's Height
hist(mdims$hgt, main="Male Height Histogram", xlab="Height",border="blue", col="green", xlim=c(150,200),probability = TRUE)
#Women's Height
hist(fdims$hgt, main="Female Height Histogram", xlab="Height",border="red", col="blue", xlim=c(145,185),probability = TRUE)
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
#Density Histogram dnorm
x <- 145:185
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")

#Q-Q Plot
qqnorm(fdims$hgt)
qqline(fdims$hgt)

#Simulate Normal Distribution & Compare 
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
qqnorm(sim_norm)
qqline(sim_norm)

#qqnormsim(fdims$hgt)

# Comparing Z table values to observed data set
1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt > 182) / length(fdims$hgt)

```
---PRACTICE---

---GRADED EXERCISES---

Q1

a. The histogram for female biiliac (pelvic) diameter (bii.di) belongs to normal probability plot letter___.

Answer = B

```{r, echo=FALSE, fig.align='center'}
qqnorm(fdims$bii.di)
qqline(fdims$bii.di)
```

b. The histogram for female elbow diameter (elb.di) belongs to normal probability plot letter ____.

Answer = C

```{r, echo=FALSE}
qqnorm(fdims$elb.di)
qqline(fdims$elb.di)
```


c. The histogram for general age (age) belongs to normal probability plot letter ____.

Answer = D

```{r, echo=FALSE, fig.align='center'}
qqnorm(bdims$age)
qqline(bdims$age)
```

d. The histogram for female chest depth (che.de) belongs to normal probability plot letter ____.

Answer = A

```{r, echo=FALSE, fig.align='center'}
qqnorm(fdims$che.de)
qqline(fdims$che.de)
```

Q2 

Note that normal probability plots C and D have a slight stepwise pattern. Why do you think this is the case?  

Answer: As noted in the study, this may be because the data isn't from a random sample, so data at the extremities my be more prone to deviate from the expected means more so than in the center.

Q3 

As you can see, normal probability plots can be used both to assess normality and visualize skewness. Make a normal probability plot for female knee diameter (kne.di). Based on this normal probability plot, is this variable left skewed, symmetric, or right skewed? Use a histogram to confirm your findings.

Answer: The data is right skewing, see plot and histogram below
```{r, echo=FALSE, fig.align='center'}
qqnorm(fdims$kne.di)
qqline(fdims$kne.di)

hist(fdims$kne.di, main="Female Knee Diameter", xlab="Height",border="red", col="grey", xlim=c(15,25),probability = TRUE)
```

