---
title: "Fav Books HTML, XML & JSON"
author: "Blandon Casenave"
date: "March 19, 2016"
output: html_document
---
Here is the code for creating a data.frame for my books from an HTML doctype.
```{r, echo=TRUE}
library(XML)
#Please create your own folder path here
booksHTML_parse <- htmlParse(file = "/Users/digitalmarketer1977/Desktop/Fav Books HTML.html" )
(booksHTML <- readHTMLTable(booksHTML_parse, header = TRUE, colClasses = NULL, skip.rows = integer(), trim = TRUE, elFun = xmlValue, as.data.frame = TRUE))

#Not clear on why a null value is thrown here, or why two lists are created.
```

Here is the code for creating a data.frame for my books from an XML doc.
```{r, echo=TRUE}
library(XML)
library(xtable)
#Please create your own folder path here
booksXML_parse <- xmlParse(file = "/Users/digitalmarketer1977/Desktop/My Fav Books.xml")
#The validation check failes on this file, but it doesn't seem to matter. Data Frame prints as expected. 
(booksXML <- xmlToDataFrame(booksXML_parse))

```

Here is the code for creating a data.frame for my books from an JSON file.

```{r, echo=TRUE}
library(XML)
require(jsonlite)
require(RJSONIO)
#check to see if JSON file is valid
isValidJSON("/Users/digitalmarketer1977/Desktop/favbooks.json")

booksJSON<- fromJSON("/Users/digitalmarketer1977/Desktop/favbooks.json", nullValue = NA, simplify = FALSE)
(booksJSONdf <-do.call("rbind", lapply (booksJSON, data.frame, stringsAsFactorsn = FALSE)))
